Supporting Information - Tackling discrepancies in freshwater invertebrate trait databases: Harmonising across continents and aggregating taxonomic resolution

### 2021-08-04 Paper/Submission_new/Latex_Supplementary_File.tex
04.08: Revisions first round completed

\subsection*{Effects of harmonisation and trait aggregation on inferences regarding trait-environment relationships}

@@

We selected the ecologically relevant traits identified by \citet{szocs_effects_2014} that increased (shredder, life cycle duration $ > 1 $ year, gills, bi/multivoltine, ovoviviparity) or decreased (univoltine, eggs in clutches/cemented or fixed, life cycle duration $ <= 1$ year) with increasing salinity and used these in the fourth corner analysis. The selected traits have also reacted to salinity in other studies (but see the discussion in \citet{szocs_effects_2014}).

For some sites, environmental and abundance data have been sampled multiple times per year. \citet{szocs_effects_2014} used the full data and did not consider repeated sampling. We applied two analysis: one where we did not consider multiple samplings. And second where we selected, for the sites with multiple samplings, only those sampling events with the highest macroinvertebrate abundances per year and accordingly only used the conductivity data for these sampling events (this was not done for the RDA approach).Fourth corner analysis was carried out in R with the \textit{fourthcorner()} function from the ade4 package.

Fourth corner analysis detected fewer trait-conductivity relationships than the RDA approach. With the original dataset significant relationships to conductivity were found for the traits shredder and ovoviviparity. In three out of the six aggregated and harmonised datasets (\textit{direct\_agg\textsubscript{mean}}, \textit{weighted\_agg}, \textit{direct\_agg\textsubscript{median}}) four significant relationships (for the traits univoltinism, bi/multivoltine, ovoviviparity, and aquatic oviposition) were detected. In the remaining three datasets (\textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and harmonised (not aggregated)) two significant relationships for the traits univoltinism and bi/multivoltinism were detected (Table \ref{stab:fc_selected_reduced}). Results were only slightly different when not considering multiple samplings (original dataset: two significant trait-conductivity relationship (bi-multivoltine, ovoviviparity), other datasets two to four, Table \ref{stab:fc_selected_full}). For all significant trait-conductivity relationships, the classification into characteristic traits for downstream and upstream sites was consistent with the results of the RDA analysis.

>>>>

\subsection*{Effects of harmonisation and trait aggregation on inferences regarding trait-environment relationships}
\label{sec:effects_harmonisation_trait_env_rel}

@@

We selected the ecologically relevant traits identified by \citet{szocs_effects_2014} that increased (shredder, life cycle duration $ > 1 $ year, gills, bi/multivoltine, ovoviviparity) or decreased (univoltine, eggs in clutches/cemented or fixed, life cycle duration $ <= 1$ year) with increasing salinity and used these in the fourth corner analysis. The selected traits also responded to salinity in other studies (but see \cite{szocs_effects_2014}).

Environmental and abundance data have been sampled multiple times per year in 12 of 15 sites. \citet{szocs_effects_2014} used the full data and ignored repeated sampling. To avoid temporal pseudoreplication, we selected, for all sites with multiple samplings only the sampling event with the highest macroinvertebrate abundance of each year and the corresponding conductivity data. Fourth corner analysis was carried out in R with the \textit{fourthcorner()} function from the ade4 package.

As expected, the fourth corner analysis detected fewer trait-conductivity relationships than the RDA approach. With the original dataset, significant relationships to conductivity were found for the traits shredder and ovoviviparity. In three out of the six aggregated and harmonised datasets (\textit{direct\_agg\textsubscript{mean}}, \textit{weighted\_agg}, \textit{direct\_agg\textsubscript{median}}) four significant relationships (for the traits univoltinism, bi/multivoltine, ovoviviparity, and aquatic oviposition) were detected. In the remaining three datasets (\textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and harmonised (not aggregated)), two significant relationships for the traits univoltinism and bi/multivoltinism were detected (Table \ref{stab:fc_selected_reduced}). The results were only slightly different when multiple samplings were ignored (original dataset: two significant trait-conductivity relationship (bi-multivoltine, ovoviviparity), other datasets two to four, Table \ref{stab:fc_selected_full}). For all significant trait-conductivity relationships, the classification into characteristic traits for downstream and upstream sites was consistent with the results of the RDA analysis.
